---
title: "Homework 1 (R)"
author: "Jinhan Cheng, jc4834@columbia.edu"
date: "October 4, 2017"
output: 
    prettydoc::html_pretty:
    theme: cayman
    highlight: github
---

```{r setup, include=FALSE}
library(knitr)
opts_chunk$set(echo = TRUE, eval = TRUE, comment="", warning = FALSE, message = FALSE, tidy.opts=list(width.cutoff=55), tidy = TRUE)
```
```{r libraries,include=FALSE}
library(prettydoc)
library(data.table)
library(plyr)
library(dplyr)
library(Hmisc)
library(DT)
library(scales)
```
```{r read file,include=FALSE}
dat <- fread(input="Homework 1 Data -- 2017.csv")
```

## Answers to Question 1:Reading the Data
* There are 2426922 rows
```{r num_rows}
nrow(x = dat)
```

* There are 13 columns
```{r num_columns}
ncol(x = dat)
```

* There are 1000 unique patients
```{r num_unique_patients}
dat[,length(unique(id))]
```
* The names of the columns are:
"**id**","**begin**","**end**","**age**",
"**diabetes**","**hypertension**","**kidney_disease**",
"**ace**","**beta.blocker**","**statin**",
"**hospital**","**heart.attack**","**death**"

* They match up with our description of the data set.
```{r colnames}
colnames(dat)
```

## Answers to Question 2:Inspection and Cleaning
* First we shall look at the types of the data by looking at the structure, it's easy to find that the types of the variables "**beta.blocker**" and "**statin**" are both "**chr**". We should convert them into numeric type.
```{r check_type}
str(dat)
```

* Checking begin
```{r check_begin}

```

* Checking end
```{r check_end}

```

* Checking age
```{r check_age}

```

* Checking diabetes
```{r check_diabetes}

```

* Checking kidney_disease
```{r check_kidney_disease}

```

* Checking ace
```{r check_ace}

```

* Checking beta.blocker
```{r check_beta_blocker}
dat[,table(beta.blocker)]
dat[,`:=`(beta.blocker,as.numeric(as.character(beta.blocker)))]
dat[,table(beta.blocker)]
```

* Checking statin
```{r check_statin}
dat[,table(statin)]
dat[,`:=`(statin,as.numeric(as.character(statin)))]
dat[,table(statin)]
```

* Checking hospital
```{r check_hospital}

```

* Checking heart.attack
```{r check_heart_attack}

```

* Checking death
```{r check_death}

```

## Answers to Question 3:Patient-Level Summaries
* The summaries at baseline for these variables are as followed:
```{r mean_clinical_factors}
baseline_dat <- dat[begin==0, ]
a_age <- baseline_dat[,round(mean(age),digits=1)]
a_diabetes <- baseline_dat[,percent(round(mean(diabetes),digits=3))] 
a_hypertension <- baseline_dat[,percent(round(mean(hypertension),digits=3))] 
a_kidney_disease <- baseline_dat[,percent(round(mean(kidney_disease),digits=3))] 
sd_age <- baseline_dat[,round(sd(age),digits=1)]
sd_diabetes <- baseline_dat[,round(sd(diabetes),digits=1)]
sd_hypertension <- baseline_dat[,round(sd(hypertension),digits=1)]
sa_kidney_disease <- baseline_dat[,round(sd(kidney_disease),digits=1)]
baseline_dat_summary <-  data.frame(variable=c("age","diabetes","hypertension","kidney disease"), average_value=c(a_age,a_diabetes,a_hypertension,a_kidney_disease), standard_deviation=c(sd_age,sd_diabetes,sd_hypertension,sa_kidney_disease)
)
datatable(baseline_dat_summary)
```

## Answers to Question 4:Counting Outcomes
* There are 49642 trips to hospital, 3262 heart attacks and 6086 deaths occurred in follow up.
```{r outcome.counts}
dat[,sum(hospital)]
dat[,sum(heart.attack)]
dat[,sum(death)]
```

## Answers to Question 5:Counting Outcomes by Medication Usage
* The results are as followed:
```{r split.outcome.counts}
Outcome_counts <- data.frame(Outcome=c("death","heart attack","hospitalization"), 
ACE_taken=c(length(which(dat[death==1,ace]==1)),length(which(dat[heart.attack==1,ace]==1)),length(which(dat[hospital==1,ace]==1))), ACE_n_taken=c(length(which(dat[death==1,ace]==0)),length(which(dat[heart.attack==1,ace]==0)),length(which(dat[hospital==1,ace]==0))), 
Beta_taken=c(length(which(dat[death==1,beta.blocker]==1)),length(which(dat[heart.attack==1,beta.blocker]==1)),length(which(dat[hospital==1,beta.blocker]==1))), Beta_n_taken=c(length(which(dat[death==1,beta.blocker]==0)),length(which(dat[heart.attack==1,beta.blocker]==0)),length(which(dat[hospital==1,beta.blocker]==0))), 
Statins_taken=c(length(which(dat[death==1,statin]==1)),length(which(dat[heart.attack==1,statin]==1)),length(which(dat[hospital==1,statin]==1))), Statins_n_taken=c(length(which(dat[death==1,statin]==0)),length(which(dat[heart.attack==1,statin]==0)),length(which(dat[hospital==1,statin]==0)))
)
datatable(Outcome_counts)
```

## Answers to Question 6:Follow up
* There are a total of 37044.5 person-years
* The average number of years of observation per patient is 3.7 years
```{r obsercation}
ob_d <- dat[,sum(end-begin)]
ob_y <- round(ob_d/365.25, digits=1)
ob_y
average_ob_y <- round(ob_y/dat[,length(unique(id))], digits=1)
average_ob_y
```

## Answers to Question 7:Utilization
* Patients spent 26958.5 years on ACE Inhibitors, 21633.6 years on Beta Blockers and 28860.3 years on Statins.
* Patients spent 10085.9 years off of ACE Inhibitors, 14276.2 years off of Beta Blockers and 8171.1 years off of Statins.
* There are 0 years missing from the observation of ACE Inhibitors, 1134.6 years missing from the observation of Beta Blockers and 13.1 years missing from the observation of Statins.
```{r utilization}
round(dat[ace==1,sum(end-begin)]/365.25,digits=1)
round(dat[beta.blocker==1,sum(end-begin)]/365.25,digits=1)
round(dat[statin==1,sum(end-begin)]/365.25,digits=1)
round(dat[ace==0,sum(end-begin)]/365.25,digits=1)
round(dat[beta.blocker==0,sum(end-begin)]/365.25,digits=1)
round(dat[statin==0,sum(end-begin)]/365.25,digits=1)
round(dat[,sum(end-begin)]/365.25-dat[ace==1,sum(end-begin)]/365.25-dat[ace==0,sum(end-begin)]/365.25,digits=1)
round(dat[,sum(end-begin)]/365.25-dat[beta.blocker==1,sum(end-begin)]/365.25-dat[beta.blocker==0,sum(end-begin)]/365.25,digits=1)
round(dat[,sum(end-begin)]/365.25-dat[statin==1,sum(end-begin)]/365.25-dat[statin==0,sum(end-begin)]/365.25,digits=1)
```

## Answers to Question 8:Crude Event Rates
* the crude rate of hospitalization is 134
* the crude rate of heart attacks is 8.8
* the crude rate of death is 16.4
```{r crude_rates}
cr_hospitalization <- round(100*dat[,sum(hospital)]/(ob_d/365.25),digits=1)
cr_hospitalization
cr_heart_attacks <- round(100*dat[,sum(heart.attack)]/(ob_d/365.25),digits=1)
cr_heart_attacks
cr_death <- round(100*dat[,sum(death)]/(ob_d/365.25),digits=1)
cr_death
cr<-data.frame(Variable=c("hospitalization","heart attacks","death"),Crude_rate=c(cr_hospitalization,cr_heart_attacks,cr_death)
)
datatable(cr)
```

## Answers to Question 9:Crude Event Rates By Medication Usage
* The results are shown in the table below

* ACE Inhibitors
```{r crude.rate.ace}
cr_hospitalization_ace_t <- round(100*dat[ace==1,sum(hospital)]/(dat[ace==1,sum(end-begin)]/365.25),digits=1)
cr_hospitalization_ace_n_t <- round(100*dat[ace==0,sum(hospital)]/(dat[ace==0,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_ace_t <- round(100*dat[ace==1,sum(heart.attack)]/(dat[ace==1,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_ace_n_t <- round(100*dat[ace==0,sum(heart.attack)]/(dat[ace==0,sum(end-begin)]/365.25),digits=1)
cr_death_ace_t <- round(100*dat[ace==1,sum(death)]/(dat[ace==1,sum(end-begin)]/365.25),digits=1)
cr_death_ace_n_t <- round(100*dat[ace==0,sum(death)]/(dat[ace==0,sum(end-begin)]/365.25),digits=1)
cr_ace <- data.frame(
Variable=c("events of hospitalization","events of heart attacks","events of death","crude rate of hospitalization","crude rate of heart attacks","crude rate of death"),
ACE_taken=c(dat[ace==1,sum(hospital)],dat[ace==1,sum(heart.attack)],dat[ace==1,sum(death)],cr_hospitalization_ace_t,cr_heart_attacks_ace_t,cr_death_ace_t),
ACE_n_taken=c(dat[ace==0,sum(hospital)],dat[ace==0,sum(heart.attack)],dat[ace==0,sum(death)],cr_hospitalization_ace_n_t,cr_heart_attacks_ace_n_t,cr_death_ace_n_t)
)
datatable(cr_ace)
```

* Beta Blockers
```{r crude.rates.beta.blocker}
cr_hospitalization_beta_t <- round(100*dat[beta.blocker==1,sum(hospital)]/(dat[beta.blocker==1,sum(end-begin)]/365.25),digits=1)
cr_hospitalization_beta_n_t <- round(100*dat[beta.blocker==0,sum(hospital)]/(dat[beta.blocker==0,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_beta_t <- round(100*dat[beta.blocker==1,sum(heart.attack)]/(dat[beta.blocker==1,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_beta_n_t <- round(100*dat[beta.blocker==0,sum(heart.attack)]/(dat[beta.blocker==0,sum(end-begin)]/365.25),digits=1)
cr_death_beta_t <- round(100*dat[beta.blocker==1,sum(death)]/(dat[beta.blocker==1,sum(end-begin)]/365.25),digits=1)
cr_death_beta_n_t <- round(100*dat[beta.blocker==0,sum(death)]/(dat[beta.blocker==0,sum(end-begin)]/365.25),digits=1)
cr_beta <- data.frame(
Variable=c("events of hospitalization","events of heart attacks","events of death","crude rate of hospitalization","crude rate of heart attacks","crude rate of death"),
BETA_taken=c(dat[beta.blocker==1,sum(hospital)],dat[beta.blocker==1,sum(heart.attack)],dat[beta.blocker==1,sum(death)],cr_hospitalization_beta_t,cr_heart_attacks_beta_t,cr_death_beta_t),
BETA_n_taken=c(dat[beta.blocker==0,sum(hospital)],dat[beta.blocker==0,sum(heart.attack)],dat[beta.blocker==0,sum(death)],cr_hospitalization_beta_n_t,cr_heart_attacks_beta_n_t,cr_death_beta_n_t)
)
datatable(cr_beta)
```

* Statins
```{r crude.rates.statin}
cr_hospitalization_statin_t <- round(100*dat[statin==1,sum(hospital)]/(dat[statin==1,sum(end-begin)]/365.25),digits=1)
cr_hospitalization_statin_n_t <- round(100*dat[statin==0,sum(hospital)]/(dat[statin==1,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_statin_t <- round(100*dat[statin==1,sum(heart.attack)]/(dat[statin==1,sum(end-begin)]/365.25),digits=1)
cr_heart_attacks_statin_n_t <- round(100*dat[statin==0,sum(heart.attack)]/(dat[statin==0,sum(end-begin)]/365.25),digits=1)
cr_death_statin_t <- round(100*dat[statin==1,sum(death)]/(dat[statin==1,sum(end-begin)]/365.25),digits=1)
cr_death_statin_n_t <- round(100*dat[statin==0,sum(death)]/(dat[statin==0,sum(end-begin)]/365.25),digits=1)
cr_statin <- data.frame(
Variable=c("events of hospitalization","events of heart attacks","events of death","crude rate of hospitalization","crude rate of heart attacks","crude rate of death"),
Statin_taken=c(dat[statin==1,sum(hospital)],dat[statin==1,sum(heart.attack)],dat[statin==1,sum(death)],cr_hospitalization_statin_t,cr_heart_attacks_statin_t,cr_death_statin_t),
Statin_n_taken=c(dat[statin==0,sum(hospital)],dat[statin==0,sum(heart.attack)],dat[statin==0,sum(death)],cr_hospitalization_statin_n_t,cr_heart_attacks_statin_n_t,cr_death_statin_n_t)
)
datatable(cr_statin)
```

## Answers to Question 10:Unadjusted Odds Ratios
* As for ACE Inhibitors:
```{r unadjusted.odds.ratio.ace}
uor_ace <- data.frame(
Variable=c("hospitalization","heart attacks","death"),
Unadjusted_odds_ratio=c(round(cr_hospitalization_ace_t/cr_hospitalization_ace_n_t,digits=2),round(cr_heart_attacks_ace_t/cr_heart_attacks_ace_n_t,digits=2),round(cr_death_ace_t/cr_death_ace_n_t,digits=2))
)
datatable(uor_ace)
```

* As for Beta Blockers:
```{r unadjusted.odds.ratio.beta.blocker}
uor_beta <- data.frame(
Variable=c("hospitalization","heart attacks","death"),
Unadjusted_odds_ratio=c(round(cr_hospitalization_beta_t/cr_hospitalization_beta_n_t,digits=2),round(cr_heart_attacks_beta_t/cr_heart_attacks_beta_n_t,digits=2),round(cr_death_beta_t/cr_death_beta_n_t,digits=2))
)
datatable(uor_beta)
```

* As for Statins:
```{r unadjusted.odds.ratio.statin}
uor_statin <- data.frame(
Variable=c("hospitalization","heart attacks","death"),
Unadjusted_odds_ratio=c(round(cr_hospitalization_statin_t/cr_hospitalization_statin_n_t,digits=2),round(cr_heart_attacks_statin_t/cr_heart_attacks_statin_n_t,digits=2),round(cr_death_statin_t/cr_death_statin_n_t,digits=2))
)
datatable(uor_statin)
```